---
title: "Fixel-wise GAM models"
output: html_notebook
---

Here we demonstrate fitting gam models at each fixel for a fixelarray dataset with  > 1000 subjects.

1. Get the package

To get the package from Github, because it's not on `CRAN` yet:

```
# go to your desired directory for the project
# cd ~/projects

# clone the repository from github
git clone https://github.com/PennBBL/FixelDB.git

# go to the repo
cd FixelDB

# specifically, use the `develop` branch
git checkout develop
```

2. Start R

In the same `FixelDB` directory, start your RStudio (recommended) or R session.

3. Load the library

Use the `devtools` package to load this library:

```{r}
# install.packages("devtools")

# devtools:load_all()

## if the above does not work, just source() the files

# source("FixelArray.R)
# source("analyse.R)
```

4. Once successful, you can use the library for your analysis:

To load fixel data, just use `FixelArray("path")`
```{r}
y <- FixelArray("/storage/fixel_stats_testing/fixel_pnc/fixel_ws_fd.h5")
y
```

Also, load your covariates table (any parsable `csv` file) with the same cohort as the h5 fixel file.

```{r}
phenos <- read.csv("./data/gam_var.csv")

head(phenos)
```

Then, you can use any of the available models to fit. There are a small handful of additional arguments you have to provide to `FixelArray` models to control efficiency options (e.g. see `?FixelArray.gam()`), but otherwise, the arguments are the passed on to the model you are fitting and so you can write the formula and formula args as normal:

```{r}

fixel_ids <- 1:1000 # note R indexing @ 1 here even though fixels start at 0

result <- FixelArray.gam(FD ~ s(Age,k=4) + sex, data = y, scalar =  "FD", phenotypes = phenos, fx=TRUE, n_cores = 12)
```

We can now see the result of the `gam` model fit at every fixel including the test statistics:

```{r}
head(result)
```

You can then write out the result with `writeResults()`

```{r}
#writeResults(y, result)
```

